---
title: "practical3.Rmd"
author: "John McIntyre"
date: "12/11/2019"
output: html_document
---

```{r, echo = FALSE}
results = "hide"
```


```{r, include = FALSE, cache = FALSE}
library(knitr)
opts_knit$set(out.format = "latex")
knit_theme$set(knit_theme$get("greyscale0"))

options(replace.assign = FALSE, width = 50)

opts_chunk$set(fig.path = "figure/graphics-",
               cache.path = "cache/graphics-",
               fig.align = "center",
               dev = "pdf", fig.width = 5, fig.height = 5,
               fig.show = "hold", cache = FALSE, par = TRUE)
knit_hooks$set(crop = hook_pdfcrop)

knit_hooks$set(par = function(before, options, envir) {
    if (before && options$fig.show != "none") {
        par(mar = c(3, 3, 2, 1), cex.lab = .95, cex.axis = .9,
            mgp = c(2, .7, 0), tcl = -.01, las = 1)
}}, crop = hook_pdfcrop)
```

```{r, echo = FALSE, include = FALSE}
library("lubridate")
```
```{r, echo = FALSE}
today()
```

# Programming: Practical 3
We wish to determine the properties on which a player is most likely to land during a game of monopoly. To simplify things, we assume there is only a single player, ignore everything to do with money and also ignore the `Get out of Jail Free Cards'.

## Monopoly: overview of the problem 
1. Begin the game on GO;
2. Roll the dice and move forward the number displayed on the dice (`current` $=$ `current` + dice roll)
3. Make a note of the new position.
    +  If we land on `Chance' or 'Community Chest', draw a card;
    +  If we land on `Go To Jail', move to Jail;
    +  If we move, make a note of the new position;
4. Go back to step 2

After rolling the dice 100,000 times or so, stop.

## Dice rolling 
When we roll a single die, each side has an equal probability of occurring. This
means we can use the `sample` function to simulate a die roll:

```{r, eval = FALSE}
sample(1:6, 1)
```

To roll two dice, we simply call this function:

```{r}
RollTwoDice = function() {
  total = sample(1:6, 1) + sample(1:6, 1)
  return(total)
}
```


#### The Monopoly board
In monopoly there are forty properties or squares, see table at the
end of this practical for a complete list. If we set the first square `Go' to be
number 1, then we can represent all forty squares as a vector in R. For example,

```{r}
## This creates a vector of 40 values;
## All values are initially zero
landings = numeric(40)
```

Then, when we land on a square we simply increase the associated `landings` entry by one. Suppose we landed on `Old Kent Rd', we would represent this as:

```{r}
landings[2] = landings[2] + 1
```
since `Old Kent Road' is square 2 (see table).

#### Going round the board
Our first go at simulating Monopoly will ignore community chest, chance cards, and the 'Go To Jail' square. This means that we are simply going round the
board. The code in the `SimulateMonopoly` function, rolls the dice a certain number of times corresponding to the value we assign to `no_of_rolls`, and stores the squares that are landed on in the vector `landings`. 

```{r}
SimulateMonopoly = function(no_of_rolls) {
  landings = numeric(40)
  ## Start at GO
  current = 1
  for (i in 1:no_of_rolls) {
    current = current + RollTwoDice()
    if (current > 40) {
      current = current - 40
    }
    landings[current] = landings[current] + 1
  }
  return(landings)
}
no_of_rolls = 50000
```

```{r, echo = FALSE}
no_of_rolls = 100
```

We can then call the function using:
```{r, warning = FALSE}
sim = SimulateMonopoly(no_of_rolls)
```

```{r, echo = FALSE}
sim = readRDS("sim1.RData")
library("tibble")
sim = as_tibble(sim)
```

We can then plot the results
```{r, include = FALSE}
library("dplyr")
```
```{r, F1, fig.keep='none', tidy=FALSE}
sim = sim %>%
  mutate(probability = value / sum(value))
library("ggplot2")
monopoly_plot = ggplot(sim, aes(x = 1:40, y = probability)) +
  geom_point() +
  labs(x = "Square", y = "Probability",
       caption = "Probability of landing on monopoly square.") +
  ylim(0, 0.06)
monopoly_plot
```

```{r, ref.label="F1", dev='pdf', fig.align="center", echo=FALSE}
```

#### Incorporating Community Chest Cards
There are three community chest squares on the board - squares 3, 18 and 34. In
the code below we will just consider square 3. There are sixteen cards in total,
hence the probability of drawing any particular card is 1/16. In the code below
we will **only implement the first two community chest cards**:

```{r}
CommunityChest = function(current) {
  goto = current
  u = runif(1)
  if (u < 1 / 16) {
    goto = 1# Move to Go
  }else if (u < 2 / 16) {
    goto = 11 #Go To Jail  :(
  }
  return(goto)
}
```

This function takes in the `current` position, with probability
1/16 we `Move to Go', with probability 1/16 we `Go to Jail' and with probability
14/16 we stay in our current position. We now alter the `simulate` function
to incorporate the `CommunityChest` function:

```{r}
SimulateMonopoly = function(no_of_rolls) {
  landings = numeric(40)
  ## Start at GO
  current = 1
  for (i in 1:no_of_rolls) {
    current = current + RollTwoDice()
    if (current > 40) {
      current = current - 40
    }
    landings[current] = landings[current] + 1
    if (current == 3) {
      cc_move = CommunityChest(current)
      if (cc_move != current) {
        current = cc_move
        landings[current] = landings[current] + 1
      }
    }
  }
  return(landings)
}
```

We can then call this function:

```{r}
sim2 = SimulateMonopoly(no_of_rolls)
```

```{r, echo = FALSE}
sim2 = readRDS("sim2.RData")
sim2 = as_tibble(sim2)
```

and plot

```{r, include = FALSE}
library("dplyr")
library("ggplot2")
```
```{r, F2, fig.keep='none', tidy=FALSE}
sim2 = sim2 %>%
  mutate(probability = value / sum(value))
library("ggplot2")
monopoly_plot2 = ggplot(sim2, aes(x = 1:40, y = probability)) +
  geom_point() +
  labs(x = "Square", y = "Probability",
       caption = "Probability of landing on monopoly square
       with the first community chest card implemented.
       Incorporating a single community chest card has very
       little effect. For this graphic, I used 2 million simulations!") +
  ylim(0, 0.06)
monopoly_plot2
```


```{r, ref.label="F2", dev='pdf', fig.align="center", echo=FALSE}
```


What is the the rationale for lines 14-20?

## Additional questions

Each question adds an additional layer of complexity to your code.

1. Add in the two other community squares, i.e. squares 18 and 34 into the `SimulateMonopoly` code.
2. Add in 'Go to Old Kent Road' into your `CommunityChest` function.
3. Square 31 is `Go To Jail.' Implement this in your main simulation function.
4. Create a `Chance` function, that implements the first six Chance
  cards. When you land on a Chance square, call this function.
5. Add in Community Chest card four.
6. Add in Chance card 8.
7. Add in Chance card 7, `Go back 3 spaces'.
8. Rolling a double (a pair of 1's, 2's, ..., 6's) is special: 
    1. Roll two dice (`total1`):  `total_score` = `total1`}
    2. If you get a double, roll again (`total2`) and `total_score` = `total1` + `total2`}
    3. If you get a double, roll again (`total3`) and `total_score` = `total1` + `total2` + `total3`}
    4. If roll three is a double, Go To Jail, otherwise move `total_score`

## Additional Information

#### Community Chest Cards
There are three community chest areas on the board (see Table). In
total, there are 16 community chest cards.

1. Advance to Go;
2. Go to jail;
3. Go to Old Kent Road;
4. Take a Chance card instead;

#### Chance Cards
A Chance card is most likely to move players. There are three chance areas on the board (see Table). There are 16 chance cards in total, of which eight cards move the player:

1. Advance to Go;
2. Advance to Trafalgar Square;
3. Advance to Pall Mall;
4. Go directly to Jail;
5. Take a trip to Marylebone Station;
6. Advance to Mayfair;
7. Go back 3 spaces;
8. Advance token to nearest Utility. The utility squares are the water works
  and the electric company.
  
| Square Number | Name | Square Number | Name |
|:------|:------|:------|:------|
| 1 | Go | 11 | Jail |
| 2 |  Old Kent Road  | 12 | Pall Mall |
| 3 |  Community Chest  | 13 | Electric Company | 
| 4 |  WhiteChapel Road  | 14 | Whitehall |
| 5 |  Income Tax  | 15 | Northumberland Avenue |
| 6 |  King's Cross Station  | 16 | Marylebone Station |
| 7 |  The Angel Islington  | 17 | Bow Street |
| 8 | Chance  | 18 | Community Chest |
| 9 |  Euston Road  | 19 | Marlborough Street |
| 10|   Pentonville Road | 20 | Vine Street |
| 21 | Free Parking | 31 | Go To Jail |
| 22 |  Strand  | 32 |  Regent Street|
| 23 |   Chance | 33 | Oxford Street | 
| 24 |   Fleet Street  | 34 | Community Chest |
| 25 |  Trafalgar Square  | 35 | Bond Street |
| 26 |   Fenchurch Street Station | 36 | Liverpool St Station |
| 27 |  Leicester Square  | 37 | Chance |
| 28 |  Coventry St | 38 | Park Lane |
| 29 |  Water Works  | 39 |  39 & Super Tax|
| 30 |   Piccadilly | 40 | Mayfair |

## Solutions

```{r, eval = FALSE}
vignette("monopoly_solutions", package = "jrProgBio")
```
