## Question 1

<!-- This question contains some maths equations, in particular two summations. If you're unsure about summations either ask your presenter or skip this question. We're not bothered about your maths theory skills, just your R skills! -->

In this question we are going to create some simple artificial DNA sequences. We will create a vector of possible bases, then we will write a loop which samples those bases, adding a new base to the end of the sequence at each iteration of the loop.

```{r results="hide", echo=TRUE}
bases = c("C", "G", "A", "T")
sequence = ""
for (i in 1:9) {
  sequence = paste0(sequence, sample(bases, 1))
}
sequence
``` 

The `for` loop above produces the artificial DNA sequence something like this: `r sequence`. Make the following changes to the above code:

1. Create an artificial DNA sequence that is 15 bases in length.
```{r}
sequence = ""
for (i in 1:15) {
  sequence = paste0(sequence, sample(bases, 1))
}
sequence
``` 

2. Create an artificial RNA sequence where thymine "T" is substituted for  uracil "U".
```{r}
bases = c("C", "G", "A", "U")
sequence = ""
for (i in 1:9) {
  sequence = paste0(sequence, sample(bases, 1))
}
sequence
``` 

3. Rewrite the loop in (2) using only the `sample()` function itself. __Hint__: you can see how sample works by checking the helpfile `?sample` and you can use `paste(collapse = "")` to concatenate them.
```{r}
paste(sample(bases, 15, replace = TRUE), collapse = "")
``` 

## Question 2

In the notes, we observed that it was straight forward to
loop through a data set and select the maximum values:
```{r echo=TRUE}
dd = data.frame(x = rnorm(10), y = rnorm(10), z = rnorm(10))

max_cols = numeric(ncol(dd))
for (i in seq_along(dd)) {
  max_cols[i] = max(dd[, i])
}
max_cols
```

 * Alter the above the code to calculate the `mean()` instead of the maximum value
 * Now, calculate the standard deviation (via `sd()`) as well as the mean. 
 You should only have a single loop!
```{r}
means = numeric(ncol(dd))
sds = numeric(ncol(dd))
for (i in seq_along(dd)) {
  means[i] = mean(dd[, i])
  sds[i] = sd(dd[, i])
}
```
    
## Solutions

Solutions are contained within this package:
```{r eval=FALSE, echo=TRUE}
vignette("solutions1", package = "jrProgBio")
```


