---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r, include = FALSE, cache = FALSE}
library(knitr)
# opts_knit$set(out.format = "latex")
knit_theme$set(knit_theme$get("greyscale0"))

options(replace.assign = FALSE, width = 50)

opts_chunk$set(fig.path = "figure/graphics-",
               cache.path = "cache/graphics-",
               fig.align = "center",
               fig.width = 4, fig.height = 4,
               fig.show = "hold", cache = FALSE, par = TRUE)
knit_hooks$set(crop = hook_pdfcrop)
```

In this question, we are going to use a `for` statement to
loop over a data set and construct some plots. The data we are going to use comes from a paper on malarial transmission traits and how they vary with temperature[^1]. This data was published to Data Dryad[^2]. To obtain the data, run the following piece of R code. 

[^1]: Shapiro LLM, Whitehead SA, Thomas MB (2017) Quantifying the effects of temperature on mosquito and parasite traits that determine the transmission potential of human malaria. PLOS Biology 15(10): e2003489. https://doi.org/10.1371/journal.pbio.2003489

[^2]: Shapiro LLM, Whitehead SA, Thomas MB (2017) Data from: Quantifying the effects of temperature on mosquito and parasite traits that determine the transmission potential of human malaria. Dryad Digital Repository. https://doi.org/10.5061/dryad.74839


```{r echo=TRUE}
data(malaria, package = "jrProgBio")
head(malaria)
``` 

The data frame `malaria` represents an experiment, where we have 6 temprerature treatments and measurements of sporozoite prevalence at over 25 days. Additionally the experiment is split into 2 blocks with each treatment in each block having 4 cups. A delightful design. We want to create a
scatter and line plot of sporozoite prevalence over time, for each treatment combination. If you decide to read the original paper, we are recreating Figure 2 from the paper but with our own code.

1. First we create a scatter plot of one treatment:

```{r F1, echo=TRUE, eval=TRUE, tidy=FALSE, message=FALSE, fig.keep="none"}
library("dplyr")
library("ggplot2")
treat_a = filter(malaria, Temperature == 21)
ggplot(treat_a, aes(x = Day, y = Sporozoite_Prevalence)) +
  geom_point()
```


2. To generate a scatter-plot for each temperature, we need to iterate over the different temperatures:

```{r results='hide', tidy=FALSE, echo = TRUE, eval=FALSE}
for (temp in unique(malaria$Temperature)) {
  group = filter(malaria, Temperature == temp)
  g = ggplot(group, aes(x = Day, y = Sporozoite_Prevalence)) +
    geom_point()
  print(g)
  readline("Hit return for next plot")
}
```
* What does `unique(malaria$Temperature)` give?
```{r}
## It gives all temperatures
```
* In the `for` loop, what variable is changing? What are it's possible values?
```{r}
## The Temperature variable is changing.
## It goes through the different temps.
```
* What does the `readline()` function do?

```{r}
## It halts execution, waits for user input
```


## Questions

1. The default axis labels aren't great. So we can change the $x$-axis label using `xlab()`:
```{r fig.keep='none', tidy=FALSE, echo=TRUE, eval=FALSE}
ggplot(group, aes(x = Day, y = Sporozoite_Prevalence)) +
  geom_point() + 
  xlab("Day")
```
```{r, fig.keep="none", eval=FALSE}
ggplot(group, aes(x = Day, y = Sporozoite_Prevalence)) +
  geom_point() + 
  xlab("Day") + 
  ylab("Sporozoite Prevalence")
```
Use the `ylab()` function to alter the $y$-axis label.

2. To add a title to a plot we could use the `ggtitle()` function, viz:

```{r F2, tidy=FALSE, fig.keep="none", eval=FALSE, echo=TRUE}
ggplot(group, aes(x = Day, y = Sporozoite_Prevalence)) +
  geom_point() + 
  xlab("Day") + 
  ylab("Sporozoite Prevalence") + 
  ggtitle("Temperature")
```

We can combine strings/characters using the `paste()` function,

```{r, eval=FALSE, echo=TRUE}
paste("Temperature", temp)
```

Rather than have a static title, make the title of each plot display the temperature.

```{r fig.keep='none', tidy=FALSE, eval=FALSE}
ggplot(group, aes(x = Day, y = Sporozoite_Prevalence)) +
  geom_point() + 
  xlab("Day") + 
  ylab("Sporozoite Prevalence") + 
  ggtitle(paste("Temperature :", temp))
```

3. The axis ranges should really be the same in all graphics. Add `ylim()` and `xlim()` to fix the range. __Hint__: Work out the range before the `for` loop using the `range()` function.

```{r  fig.keep='none', tidy=FALSE, eval=FALSE}
ylims = range(malaria$Sporozoite_Prevalence)
xlims = range(malaria$Day)

ggplot(group, aes(x = Day, y = Sporozoite_Prevalence)) +
  geom_point() + 
  xlab("Day") + 
  ylab("Sporozoite Prevalence") + 
  ggtitle(paste("Temperature :", temp)) + 
  xlim(xlims[1], xlims[2]) + 
  ylim(ylims[1], ylims[2])
```

4. For each graph, we want to add colour to the points to represent which experimental block they come from. 

```{r, fig.keep='none', tidy=FALSE, eval=FALSE}
ggplot(group, aes(x = Day, y = Sporozoite_Prevalence,
                  colour = factor(Block))) +
  geom_point() + 
  xlab("Day") + 
  ylab("Sporozoite Prevalence") + 
  ggtitle(paste("Temperature :", temp)) + 
  xlim(xlims[1], xlims[2]) + 
  ylim(ylims[1], ylims[2])
```


5. For extra bonus points see if you can add the lines for each cup. You should end up with two colours but 4 line traces of each colour (and your points of course). Use the `geom_line()` function in the same way you used the `geom_point()` function.

__Hint__: For the `geom_line()` have a look at the `group` aesthetic and the `interaction()` function.

```{r, fig.keep='none', tidy=FALSE, eval=FALSE}
ggplot(group, aes(x = Day, y = Sporozoite_Prevalence,
                  colour = factor(Block))) +
  geom_point() + 
  geom_line(aes(group = interaction(Block,Cup))) + 
  xlab("Day") + 
  ylab("Sporozoite Prevalence") + 
  ggtitle(paste("Temperature :", temp)) + 
  xlim(xlims[1], xlims[2]) + 
  ylim(ylims[1], ylims[2])
```



You should end up with plots that looks something like this for each temperature:

```{r message=FALSE, tidy=FALSE, echo=FALSE, eval=TRUE, fig.height=3}
temp = 34
group = filter(malaria, Temperature == temp)
ylims = range(malaria$Sporozoite_Prevalence)
xlims = range(malaria$Day)
ggplot(group, aes(x = Day, y = Sporozoite_Prevalence,
                  colour = factor(Block))) +
  geom_point() + 
  geom_line(aes(group = interaction(Block,Cup))) + 
  xlab("Day") + 
  ylab("Sporozoite Prevalence") + 
  ggtitle(paste("Temperature :", temp)) + 
  xlim(xlims[1], xlims[2]) + 
  ylim(ylims[1], ylims[2])
```

Can you think of further ways to use programming to make this code neater?

6. Suppose we wanted to save these graphs in a pdf file. Add the `pdf()` function to your code save the resulting graph:

```{r, eval=FALSE, echo=TRUE}
# decide on a filename and path
filename = "my_awesome_figure.pdf"

# do your plotting

# close the connection to the file
dev.off()
```

7. Finally put your code, i.e. the `for` loop and plotting commands, in a function which takes the data frame as an argument.

## Solutions

Solutions are contained within this package:
```{r eval=FALSE, echo=TRUE}
vignette("solutions2", package = "jrProgBio")
```

```{r message=FALSE, tidy=FALSE, eval=FALSE}
## FULL SOLUTION
viewgraphs = function(malaria, save=FALSE, filename = "graphs.pdf") {

  if (save) {
    pdf(filename)
  }

  ylims = range(malaria$Sporozoite_Prevalence)
  xlims = range(malaria$Day)

  for (temp in unique(malaria$Temperature)) {

    group = filter(malaria, Temperature == temp)

    g = ggplot(group, aes(x = Day, y = Sporozoite_Prevalence,
                          colour = factor(Block))) +
      geom_point() + 
      geom_line(aes(group = interaction(Block,Cup))) + 
      xlab("Day") + 
      ylab("Sporozoite Prevalence") + 
      ggtitle(paste("Temperature :", temp)) + 
      xlim(xlims[1], xlims[2]) + 
      ylim(ylims[1], ylims[2])
    print(g)
  }

  if (save) {
    dev.off()
  }

}
```
