```{r, include = FALSE, cache = FALSE}
library(knitr)
# opts_knit$set(out.format = "latex")
knit_theme$set(knit_theme$get("greyscale0"))

options(replace.assign=FALSE,width=50)

opts_chunk$set(fig.path='figure/graphics-', 
               cache.path='cache/graphics-', 
               fig.align='center', 
               fig.width=4, fig.height=4, 
               fig.show='hold', cache=FALSE, par=TRUE)
knit_hooks$set(crop=hook_pdfcrop)
```

In this question, we are going to use a `for` statement to
loop over a data set and construct some plots. The data we are going to use comes from a paper on malarial transmission traits and how they vary with temperature[^1]. This data was published to Data Dryad[^2]. To obtain the data, run the following piece of R code. 

[^1]: Shapiro LLM, Whitehead SA, Thomas MB (2017) Quantifying the effects of temperature on mosquito and parasite traits that determine the transmission potential of human malaria. PLOS Biology 15(10): e2003489. https://doi.org/10.1371/journal.pbio.2003489

[^2]: Shapiro LLM, Whitehead SA, Thomas MB (2017) Data from: Quantifying the effects of temperature on mosquito and parasite traits that determine the transmission potential of human malaria. Dryad Digital Repository. https://doi.org/10.5061/dryad.74839


```{r echo=TRUE}
data(malaria, package = "jrProgBio")
head(malaria)
``` 

The data frame `malaria` represents an experiment, where we have 6 temprerature treatments and measurements of sporozoite prevalence at over 25 days. Additionally the experiment is split into 2 blocks with each treatment in each block having 4 cups. A delightful design. We want to create a
scatter and line plot of sporozoite prevalence over time, for each treatment combination. If you decide to read the original paper, we are recreating Figure 2 from the paper but with our own code.

1. First we create a scatter plot of one treatment:

```{r F1, echo=TRUE, eval=TRUE, tidy=FALSE, message=FALSE, fig.keep="none"}
library(dplyr)
treat_a = filter(malaria, Temperature == 21)
plot(treat_a$Sporozoite.Prevalence, treat_a$Day)
```


1. To generate a scatter-plot for each temperature, we need to iterate over the different temperatures:

```{r results='hide', fig.keep='none', tidy=FALSE, echo = TRUE, eval=FALSE}
for(temp in unique(malaria$Temperature)) {
  group = filter(malaria, Temperature == temp)
  plot(group$Sporozoite.Prevalence, group$Day)
  readline("Hit return for next plot")
}
```
* What does `unique(malaria$Temperature)` give?
```{r}
## It gives all temperatures
```
* In the `for` loop, what variable is changing? What are it's possible values?
```{r}
## The Temperature variable is changing.
## It goes through the different temps.
```
* What does the `readline()` function do?

```{r}
## It halts execution, waits for user input
```


## Questions

1. The default axis labels aren't great. So we can change the $x$-axis label using `xlab`:
```{r fig.keep='none', tidy=FALSE, echo=TRUE}
plot(group$Sporozoite.Prevalence, group$Day, xlab = "Day")
```
```{r, fig.keep="none"}
plot(group$Sporozoite.Prevalence, group$Day,
     xlab = "Day", ylab = "Sporozoite Prevalence")
```
Use the `ylab` argument to alter the $y$-axis label.

1. To add a title to a plot we use the `main` argument, viz:

```{r F2, tidy=FALSE, fig.keep="none"}
plot(group$Sporozoite.Prevalence, group$Day,
     xlab = "Day", ylab = "Sporozoite Prevalence",
     main = "Treatment")
```

We can combine strings/characters using the `paste()` function,

```{r, eval=FALSE}
paste("Treatment", temp)
```

Rather than have a static title, make the title of each plot display the treatment type.

```{r fig.keep='none', tidy=FALSE}
plot(group$Sporozoite.Prevalence, group$Day,
     xlab = "Day", ylab = "Sporozoite Prevalence",
     main = paste("Temperature :", temp))
```

1. The y-axis range should really be the same in all graphics. Add a `ylim` argument
to fix the range. __Hint__: Work out the range before the `for` loop.

```{r  fig.keep='none', tidy=FALSE}
range(malaria$Sporozoite.Prevalence)
plot(group$Sporozoite.Prevalence, group$Day,
     xlab = "Day", ylab = "Sporozoite Prevalence",
     main = paste("Temperature :", temp),
     ylim = c(0, 1))
```

1. At each iteration, use the `message()` function to print the average sporoziote prevalence across all time points and replicates.

```{r results='hide', message=FALSE}
##Within the for loop have the line
message(mean(group$Sporozoite.Prevalence))
```

<!-- 1. On each graph, highlight any observations with a blue point if they are -->
<!-- larger than the mean + standard deviations or less than the mean -  -->
<!-- standard deviations. Use the `points()` function to highlight a point. -->
<!-- __Hint__: You don't need `if` statements here.  -->
<!-- Just subset your data frame and pass this new data frame to the points function. -->
<!-- For example, to highlight the points (1,2) and (3, 4) we use the command: -->

<!-- ```{r echo=-1, fig.keep='none'} -->
<!-- plot(0) -->
<!-- points(c(1, 3), c(2, 4), col=2) -->
<!-- ``` -->

<!-- ```{r fig.keep='none', message=FALSE, tidy=FALSE} -->
<!-- plot(group$time, group$values, -->
<!--      ylab = "Measurement", xlab="Time", -->
<!--      main=paste("Treatment", treatment)) -->

<!-- ##Calculate the limits -->
<!-- values = group$values -->
<!-- message(mean(values)) -->
<!-- upper_lim = mean(values) + sd(values) -->
<!-- lower_lim = mean(values) - sd(values) -->

<!-- ##Extract the points -->
<!-- outliers = filter(group, values > upper_lim | values < lower_lim) -->
<!-- ##pch=19 gives a solid dot -->
<!-- ##See ?points -->
<!-- points(outliers$time, outliers$values, col=4, pch=19) -->
<!-- ``` -->


<!-- 1. Suppose we wanted to save individual graphs in a pdf file. Add the `pdf()` -->
<!-- function to your code save the resulting graph. To get unique file names, use the paste -->
<!-- command: -->

<!-- ```{r} -->
<!-- filename = paste0("file", treatment, ".pdf") -->
<!-- ``` -->
<!-- 1. Put your code, i.e. the `for` loop and plotting commands, in a function which takes the data frame as an argument. -->


<!-- ## Solutions -->

<!-- Solutions are contained within this package: -->
<!-- ```{r eval=FALSE, echo=TRUE} -->
<!-- vignette("solutions2", package = "jrProgramming") -->
<!-- ``` -->


<!-- ```{r tidy=FALSE} -->
<!-- ## FULL SOLUTION -->
<!-- viewgraphs = function(exper,  save=FALSE) { -->
<!--   for(treat in unique(exper$treat)) { -->
<!--     if(save) { -->
<!--       filename = paste("file", treat, ".pdf", sep="") -->
<!--       pdf(filename) -->
<!--     } -->

<!--     plot(group$time, group$values, -->
<!--          ylab="Measurement", xlab="Time", -->
<!--          main=paste("Treatment", treat)) -->

<!--     ##Calculate the limits -->
<!--     values = group$values -->
<!--     message(mean(values)) -->
<!--     upper_lim = mean(values) + sd(values) -->
<!--     lower_lim = mean(values) - sd(values) -->

<!--     ##Extract the points -->
<!--     outliers = filter(group, values > upper_lim | values < lower_lim) -->
<!--     ##pch=19 gives a solid dot -->
<!--     ##See ?points -->
<!--     points(outliers$time, outliers$values, col=4, pch=19) -->
<!--     if(save){ -->
<!--       dev.off() -->
<!--     } else {   -->
<!--       readline("Hit return for next plot\n") -->
<!--     } -->
<!--   } -->
<!-- } -->
<!-- ``` -->
