---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r, include = FALSE, cache = FALSE}
library(knitr)
# opts_knit$set(out.format = "latex")
knit_theme$set(knit_theme$get("greyscale0"))

options(replace.assign=FALSE,width=50)

opts_chunk$set(fig.path='figure/graphics-', 
               cache.path='cache/graphics-', 
               fig.align='center', 
               fig.width=4, fig.height=4, 
               fig.show='hold', cache=FALSE, par=TRUE)
knit_hooks$set(crop=hook_pdfcrop)
```

In this question, we are going to use a `for` statement to
loop over a data set and construct some plots. The data we are going to use comes from a paper on malarial transmission traits and how they vary with temperature[^1]. This data was published to Data Dryad[^2]. To obtain the data, run the following piece of R code. 

[^1]: Shapiro LLM, Whitehead SA, Thomas MB (2017) Quantifying the effects of temperature on mosquito and parasite traits that determine the transmission potential of human malaria. PLOS Biology 15(10): e2003489. https://doi.org/10.1371/journal.pbio.2003489

[^2]: Shapiro LLM, Whitehead SA, Thomas MB (2017) Data from: Quantifying the effects of temperature on mosquito and parasite traits that determine the transmission potential of human malaria. Dryad Digital Repository. https://doi.org/10.5061/dryad.74839


```{r echo=TRUE}
data(malaria, package = "jrProgBio")
head(malaria)
``` 

The data frame `malaria` represents an experiment, where we have 6 temprerature treatments and measurements of sporozoite prevalence at over 25 days. Additionally the experiment is split into 2 blocks with each treatment in each block having 4 cups. A delightful design. We want to create a
scatter and line plot of sporozoite prevalence over time, for each treatment combination. If you decide to read the original paper, we are recreating Figure 2 from the paper but with our own code.

1. First we create a scatter plot of one treatment:

```{r F1, echo=TRUE, eval=TRUE, tidy=FALSE, message=FALSE, fig.keep="none"}
library(dplyr)
treat_a = filter(malaria, Temperature == 21)
plot(treat_a$Sporozoite.Prevalence, treat_a$Day)
```


1. To generate a scatter-plot for each temperature, we need to iterate over the different temperatures:

```{r results='hide', tidy=FALSE, echo = TRUE, eval=FALSE}
par(mfrow = c(2, 3))
for(temp in unique(malaria$Temperature)) {
  group = filter(malaria, Temperature == temp)
  plot(group$Day, group$Sporozoite.Prevalence)
}
```
* What does `unique(malaria$Temperature)` give?
```{r}
## It gives all temperatures
```
* In the `for` loop, what variable is changing? What are it's possible values?
```{r}
## The Temperature variable is changing.
## It goes through the different temps.
```
* What does the `par()` function do? And how can you use the `mfrow` argument?

```{r}
## It sets up the graphical parameters of the plot and the mfrow argumement is used to split the plotting area into panels.
```


## Questions

1. The default axis labels aren't great. So we can change the $x$-axis label using `xlab`:
```{r fig.keep='none', tidy=FALSE, echo=TRUE, eval=FALSE}
plot(group$Day, group$Sporozoite.Prevalence, xlab = "Day")
```
```{r, fig.keep="none", eval=FALSE}
plot(group$Day, group$Sporozoite.Prevalence,
     xlab = "Day", ylab = "Sporozoite Prevalence")
```
Use the `ylab` argument to alter the $y$-axis label.

1. To add a title to a plot we use the `main` argument, viz:

```{r F2, tidy=FALSE, fig.keep="none", eval=FALSE}
plot(group$Day, group$Sporozoite.Prevalence,
     xlab = "Day", ylab = "Sporozoite Prevalence",
     main = "Treatment")
```

We can combine strings/characters using the `paste()` function,

```{r, eval=FALSE}
paste("Treatment", temp)
```

Rather than have a static title, make the title of each plot display the treatment type.

```{r fig.keep='none', tidy=FALSE, eval=FALSE}
plot(group$Day, group$Sporozoite.Prevalence,
     xlab = "Day", ylab = "Sporozoite Prevalence",
     main = paste("Temperature :", temp))
```

1. The axis ranges should really be the same in all graphics. Add `ylim` and `xlim` arguments to fix the range. __Hint__: Work out the range before the `for` loop using the `range()` function.

```{r  fig.keep='none', tidy=FALSE, eval=FALSE}
ylims = range(malaria$Sporozoite.Prevalence)
xlims = range(malaria$Day)

plot(group$Day, group$Sporozoite.Prevalence,
     xlab = "Day", ylab = "Sporozoite Prevalence",
     main = paste("Temperature :", temp),
     ylim = ylims, xlim = xlims)
```

1. For each graph, we want to add colour to the points to represent which experimental block they come from. We can first use `type = "n"` to produce and empty plot, then create additionl subsets of the data and use the `points()` function to add points to the graph.

__Hint__: You don't need `if` statements here.
Just subset your data frame and pass this new data frame to the points function.
Below is an example of how we might highlight the 5th and 6th points in a dummy dataset:

```{r echo=-1, fig.keep='none', eval=FALSE}
# create the data
x = 1:10
y = rnorm(10)

# plot it
plot(x, y)

# add a few coloured points
points(x[5], y[5], col="red")
```

```{r fig.keep='none', message=FALSE, tidy=FALSE, eval=FALSE}
plot(group$Day, group$Sporozoite.Prevalence,
     xlab = "Day", ylab = "Sporozoite Prevalence",
     main = paste("Temperature :", temp),
     ylim = ylims, xlim = xlims, type = "n")

##Extract the points
block_0 = filter(group, Block == 0)
block_1 = filter(group, Block == 1)

##pch=19 gives a solid dot
##See ?points
points(block_0$Day, block_0$Sporozoite.Prevalence, col="blue", pch=19)
points(block_1$Day, block_1$Sporozoite.Prevalence, col="red", pch=19)
```

1. For extra bonus points see if you can add an internal loop for each block to add the lines for each cup. You should end up with two colours but 4 line traces of each colour (and your points of course). Use the `lines()` function in the same way you used the `points()` function.

__Hint__: This is an analogous loop to the one we started with, for each cup first filter the data down and then add some lines. You are allowed to have as deeply nested loops as you like.

```{r fig.keep='none', message=FALSE, tidy=FALSE, eval=FALSE}
plot(group$Day, group$Sporozoite.Prevalence,
     xlab = "Day", ylab = "Sporozoite Prevalence",
     main = paste("Temperature :", temp),
     ylim = ylims, xlim = xlims, type = "n")

##Extract the points
block_0 = filter(group, Block == 0)
block_1 = filter(group, Block == 1)

##pch=19 gives a solid dot
##See ?points
points(block_0$Day, block_0$Sporozoite.Prevalence, col="blue", pch=19)
for(cup in unique(block_0$Cup)) {
  cup_group = filter(block_0, Cup == cup)
  lines(cup_group$Day, cup_group$Sporozoite.Prevalence, col = "blue")
}
points(block_1$Day, block_1$Sporozoite.Prevalence, col="red", pch=19)
for(cup in unique(block_1$Cup)) {
  cup_group = filter(block_1, Cup == cup)
  lines(cup_group$Day, cup_group$Sporozoite.Prevalence, col = "red")
}
```

You should end up with a plot that looks something like this:

```{r message=FALSE, tidy=FALSE, echo=FALSE, eval=TRUE, fig.height=3}
par(mfrow = c(2, 3), mar = c(2.5, 2.5, 1, 1), mgp = c(1, 0.5, 0))

ylims = range(malaria$Sporozoite.Prevalence)
xlims = range(malaria$Day)

for(temp in unique(malaria$Temperature)) {
  
  group = filter(malaria, Temperature == temp)
  
  cex_val = 0.6
  plot(group$Day, group$Sporozoite.Prevalence,
       xlab = "Day", ylab = "Sporozoite Prevalence",
       main = paste("Temperature :", temp),
       ylim = ylims, xlim = xlims, type = "n",
       cex.lab=cex_val, cex.axis=cex_val, cex.main=cex_val, cex.sub=cex_val)
  
  ## Extract the points
  block_0 = filter(group, Block == 0)
  block_1 = filter(group, Block == 1)
  
  ## Plot the subsets
  blue = "#0000ff75"
  points(block_0$Day, block_0$Sporozoite.Prevalence, 
         col=blue, pch=19, cex = cex_val)
  for(cup in unique(block_0$Cup)) {
    cup_group = filter(block_0, Cup == cup)
    lines(cup_group$Day, cup_group$Sporozoite.Prevalence, 
          col = blue, cex = cex_val)
  }
  
  red = "#ff000075"
  points(block_1$Day, block_1$Sporozoite.Prevalence, 
         col=red, pch=19, cex = cex_val)
  for(cup in unique(block_1$Cup)) {
    cup_group = filter(block_1, Cup == cup)
    lines(cup_group$Day, cup_group$Sporozoite.Prevalence, 
          col = red, cex = cex_val)
  }
  
}
```

Can you think of further ways to use programming to make this code neater?
1. Suppose we wanted to save this multi-panel graph in a pdf file. Add the `pdf()` function to your code save the resulting graph:

```{r, eval=FALSE, echo=TRUE}
# decide on a filename and path
filename = "my_awesome_figure.pdf"

# do your plotting
plot(0)

# close the connection to the file
dev.off()
```

1. Finally put your code, i.e. the `for` loop and plotting commands, in a function which takes the data frame as an argument.

## Solutions

Solutions are contained within this package:
```{r eval=FALSE, echo=TRUE}
vignette("solutions2", package = "jrProgBio")
```

```{r message=FALSE, tidy=FALSE, eval=FALSE}
## FULL SOLUTION
viewgraphs = function(malaria, save=FALSE) {
  
  if(save) {
    filename = paste("file", treat, ".pdf", sep="")
    pdf(filename)
  }
  
  par(mfrow = c(2, 3))
  
  ylims = range(malaria$Sporozoite.Prevalence)
  xlims = range(malaria$Day)
  
  for(temp in unique(malaria$Temperature)) {
    
    group = filter(malaria, Temperature == temp)
    
    plot(group$Day, group$Sporozoite.Prevalence,
         xlab = "Day", ylab = "Sporozoite Prevalence",
         main = paste("Temperature :", temp),
         ylim = ylims, xlim = xlims, type = "n")
    
    ## Extract the points
    block_0 = filter(group, Block == 0)
    block_1 = filter(group, Block == 1)
    
    ## Plot the subsets
    points(block_0$Day, block_0$Sporozoite.Prevalence, col="blue", pch=19)
    for(cup in unique(block_0$Cup)) {
      cup_group = filter(block_0, Cup == cup)
      lines(cup_group$Day, cup_group$Sporozoite.Prevalence, col = "blue")
    }
    
    points(block_1$Day, block_1$Sporozoite.Prevalence, col="red", pch=19)
    for(cup in unique(block_1$Cup)) {
      cup_group = filter(block_1, Cup == cup)
      lines(cup_group$Day, cup_group$Sporozoite.Prevalence, col = "red")
    }
    
  }
  
  if(save){
    dev.off()
  }
  
}
```
