## Question 1

Consider the following simple function that produces a DNA sequence:
```{r echo=TRUE}
n = 5
sequence = function() {
  bases = c("C", "G", "A", "T")
  n = 10
  paste(sample(bases, n, TRUE), collapse = "")
}
sequence()
```

1. Why does the final line return a sequence of length 10 and not 5.
```{r }
## sequence uses the local variable n
```

2. Delete line 4 in the above piece of code. Now change `sequence()` to allow `v` to be passed as an argument, i.e. we can write `sequence(5)`. Call this function to make sure it works.

```{r}
sequence = function(n) {
  bases = c("C", "G", "A", "T")
  paste(sample(bases, n, TRUE), collapse = "")
}
sequence(5)
```


## Question 2

```{r echo=TRUE}
one = function(x = 10) {
  return(x)
}

two = function(x) {
  return(x)
}
``` 

1. Why does 
```{r echo=TRUE}
one()
```
work, but this raises an error
```{r eval=FALSE, echo=TRUE}
two()
```

```{r}
## two() expects an argument x, but
## we haven't given one and there is
## no default.
```

2. Change the `sequence` function so that it has a default value for `n`.

```{r }
sequence = function(n = 10) {
  bases = c("C", "G", "A", "T")
  paste(sample(bases, n, TRUE), collapse = "")
}
sequence()
``` 
    
## Question 3

So far we have been creating random DNA sequences. Lets do another routine job, split them into codons. You could approach this in a number of ways, but lets start with something like this (you will need the `stringr` package).

```{r, echo=TRUE}
library(stringr)
# create the sequence
s = sequence()

# establish indeces of the first and last base in the codon
offset = 0
start = seq(1 + offset, str_length(s), by = 3)
end = start + 2

# retrieve the codons
str_sub(s, start, end)
```

1. In the code above, delete the first two lines and put the code into a function called `codons`.

```{r}
codons = function(sequence) {
  offset = 0
  start = seq(1 + offset, nchar(sequence), by = 3)
  end = start + 2
  str_sub(sequence, start, end)
}
```

2. Alter the function such that offset is an argument with the default value of 0.

```{r}
codons = function(sequence, offset = 0) {
  start = seq(1 + offset, nchar(sequence), by = 3)
  end = start + 2
  str_sub(sequence, start, end)
}
```

3. Lets imagine that codons could be any length, alter the function such that the length of the codons is an argument of the function with a default value.

```{r}
codons = function(sequence, offset = 0, length = 3) {
  start = seq(1 + offset, nchar(sequence), by = length)
  end = start + length - 1
  str_sub(sequence, start, end)
}
```

    
## Solutions

Solutions are contained within this package:
```{r eval=FALSE, echo=TRUE}
vignette("solutions2", package = "jrProgBio")
```
